---
title: "Data Analaysis with the Arctic shorebird csv files"
format: html
---


```{r, message=FALSE}
library(tidyverse)
library(lubridate)
library(RColorBrewer)
```


## Define path and filenames

```{r}
path_raw <- "data/raw"

sites_csv <- "site.csv"

snowcover_csv <- "snow_cover.csv"

nests_csv <- "ASDN_Bird_nests.csv"

eggs_csv <- "ASDN_Bird_eggs.csv"

species_csv <- "species.csv"

lemmings_csv <- "ASDN_Daily_pred_lemm.csv"
```


## Question 1: Does the snow amount & timing impact the number of eggs? Is it species dependant?

### Summarize the nests table

Read the nest data in

```{r}
nests_data <- read_csv(file.path(path_raw, nests_csv), 
                       show_col_types = FALSE)

glimpse(nests_data)
```

Yeah! `Date_found` is already in date format! We can thus directly aggregate the data by months:


```{r}
# we want the number of nests 
nests_month <- nests_data %>% 
  mutate(month = lubridate::month(Date_found)) %>% 
  group_by(Species, Site, Year, month) %>% 
  # no nests = no. rows for each group
  summarise(n_nests = n())
  
nests_month
```

### Summarize the snow cover table

Read the snow cover data in

```{r}
snowcover_data <- read_csv(file.path(path_raw, snowcover_csv),
                           show_col_types = FALSE)

glimpse(snowcover_data)
```

The Date is stored as a character we thus first we need to transform the date in date format before computing the monthly aggregate:

```{r}
snowcover_data_fixed <- snowcover_data %>% 
  # this gives us a warning that 72 rows could not have this function work properly
  # we need to investigate
  # if it fails, it goes to NA
  mutate(Date2 = lubridate::as_date(Date))
```

We have 72 errors. Let's have a look at the date which failed:

```{r}
# peeping we see Date "8&9 june 06"
snowcover_data_fixed %>% 
  filter(is.na(Date2))
```


It is because 2 days were entered as one value `8&9 june 06`, it is not in the expected format

```{r}
# this overwrites the NA values in the Date2 column to match what they should have been when running originally, if they'd been cleaned properly
snowcover_data_fixed <-snowcover_data_fixed %>% 
  mutate(Date2 = case_when(
    Date == "8&9 june 06" ~ "09-June-06",
    .default = Date
  )) %>% 
  mutate(Date2 = as_date(Date, format = "%d-%b-%y"))
# peeping we see Date "8&9 june 06"


snowcover_data_fixed %>% 
  filter(is.na(Date2))

snowcover_data_fixed
```


```{r}
snowcover_data_fixed <-snowcover_data_fixed %>% 
  mutate(Date = case_when(
    Date == "8&9 june 06" ~ "09-June-06",
    .default = Date
  )) %>% 
  mutate(Date2 = as_date(Date, format = "%d-%b-%y"))

snowcover_data_fixed
```

Mmm the Dates are not what we were expecting... It generally means there are different date types. Let's start by fixing the problematic dates before transforming it to the date format.

```{r}

```


::: {.callout-note}
## Fun fact
You might be wondering why I started with `as_date()` instead of directly using `dmy()`. Well turns out that while preparing this exercise I discovered a bug in this `lubridate` function!! See <https://github.com/tidyverse/lubridate/issues/1161>
:::


Now let's compute the monthly statistics

cummulative snow days, mean snow cover, and group_by site, year, month

```{r}
# adding the snow days with a threshold at 10%
snowcover_data_fixed <- snowcover_data_fixed %>% 
  
  mutate(month = month(Date2)) %>% 

  group_by(Site, Year, month) %>% 
  
  mutate(snow_days = Snow_cover > 10) %>% 
  
  summarize(n_snowdays = sum(snow_days, na.rm = TRUE),
            avg_snow_cover = mean(Snow_cover, na.rm = TRUE))

snowcover_data_fixed
```

 
### Join the snow cover and nest tables

```{r}
combined_tables <- inner_join(nests_month, snowcover_data_fixed,
                              join_by(Site, month, Year))

combined_tables
```

Well in the end we did not have much information by months (mainly 2 months per year focused on June and July), so let's aggregate by year instead:

```{r}
combined_years <- combined_tables %>% 
  group_by(Species, Site, Year) %>% 
  summarize(n_snowdays = sum(n_snowdays, na.rm = TRUE),
            avg_snow_cover = mean(avg_snow_cover, na.rm = TRUE),
            n_nests = sum(n_nests, na.rm = TRUE)
            )
  
combined_years
```

### Plots

```{r}
ggplot(combined_years, aes(x = n_snowdays, y = n_nests, color = Species)) +
  geom_point() +
  scale_color_brewer(palette = "Set3")
```

```{r}
library(here)
common_names <- read_csv(here("data", "raw", "species.csv"),
                         show_col_types = FALSE) 
common_names
```
```{r}
species_year <- combined_years %>% 
  left_join(common_names, join_by(Species == Code))

species_year
```
```{r}
ggplot(species_year, aes(x = n_snowdays, y = n_nests, color = Common_name)) +
  geom_point() +
  scale_color_brewer(palette = "Set3")
```

What do you think is the answer to our question? 


## Question 2: Which shorebird species makes the most eggs?

### Import the data

```{r}
sites_data <- read_csv(file.path(path_raw, sites_csv))

eggs_data <- read_csv(file.path(path_raw, eggs_csv))

species_data <- read_csv(file.path(path_raw, species_csv))
```

Let's compute the number of eggs per nest. The column `Egg-num` seems to be a good candidate to estimate the number of eggs per nest:

```{r}
eggs_sum <- 
```

Join with the nest data:

```{r}

```

Group by species:

```{r}

```

OK we might want to have more explicit names for the species!, Luckily we have the species table to help with that:

```{r}

```

Some cleaning of the data frame to put the average at the end and remove the Relevance column:

```{r}

```

